---
title: "Data 607 Assignment 2"
output: html_document
#params:
#  pwd:
#    label: "Enter the database password, please!"
#    value: ""
#    input: password
---
#### ***Rathish Parayil Sasidharan***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview

Created a survey to rate six movies **GREENLAND,THE CROODS: A NEW AGE,MULAN,SUMMERLAND,GREYHOUND,THE GENTLEMEN** \
 with rating Excellent (5),Good (4),Not Bad(3), Bad(2),Awful(1)  and collected the survey response.
 
[Please find the survey ](https://www.surveymonkey.com/r/F82FVCV)

High level tasks performed as part of this assignment

1. Created a survey in surveymonkey and collected the response \
2. Created AWS RDS instance and configured it to access using public url \
3. Created 2 normalized tables,one for movie and other for movie rating \
4. Inserted the response data into database table using a script \
5. In order to protect the data ,created a readonly user account on the db \ 
6. Read the data from database from the R Markup \
7. Eliminated N/A data \
8. Data transformation and Visualization



## Prerequisites
Load required packages
```{r message=FALSE}
#install.packages("RMySQL")
library("RMySQL")
library("dplyr")
library("ggplot2")
library("tidyr")
#library("config")
 
```

## Loading the dataframe from the Data base


```{r}

# Create a connection Object to MySQL database.

#password can also be read as a user input by knit with parameter option
#or can also be read from a config file(yaml)
#dw <- config::get("applicationdb")
#print(dw$server)
#print(dw$uid)
#print(dw$pwd)
#mysqlconnection = dbConnect(MySQL(), user = dw$uid, password = dw$pwd, dbname = 'database_name',
 #  host = dw$server)



mysqlconnection = dbConnect(MySQL(), user = 'readOnly', password = 'welcome', dbname = 'database_name',
   host = 'rassnw2.c4h7w2meulft.us-east-1.rds.amazonaws.com')

# List the tables available in this database.
 #dbListTables(mysqlconnection)
 
 # Query the movie rating tables to get all rating and associated movies.
 result = dbSendQuery(mysqlconnection, "select mr.id,mv.movie_name,mr.rating from movie_rating mr join movie mv on mr.movie_id = mv.movie_id")

# Store the result in a R data frame object. n = 30 is used to fetch first 30 rows.
employeedata = fetch(result, n = 30)
print(employeedata)

```


### Clean up the data
```{r}
employee_valid_data <- employeedata %>% drop_na()

# Conver the ratings into a Readable Text
 employee_valid_data <- mutate(employee_valid_data,review = case_when(rating == 5 ~'Excellent',rating == 4 ~'Good',rating == 3 ~'Not Bad',rating == 2 ~'Bad',rating == 1 ~'Awful' ))
 
 print(employee_valid_data)
 
```


### Data visualization
How each movies are rated by the users ?\
Plot a graph with list of movies vs number of rating for each category

```{r}
ggplot(count(employee_valid_data,movie_name,review), aes(fill=review, y=n, x=review)) +
     geom_bar(position="dodge", stat="identity") +
     ggtitle("Movie rating") +
     facet_wrap(~movie_name)+xlab("")+ylab("Number of Rating") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### Conclusion

According to the survey result "Greenland" movie has got better rating

