---
title: "Data607 Project1"
output: html_document
---
#### ***Rathish Parayil Sasidharan***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this project, you’re given a text file with chess tournament results where the information has some structure.  Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents 



### Prerequisites
Load required packages
```{r}
library("stringr")

```


## Read the input text file
```{r}

tournamentDf<-read.csv("https://raw.githubusercontent.com/rathish-ps/Data607-Assignment/main/tournamentinfo.txt",header=FALSE)

tournamentData <- matrix(unlist(tournamentDf), byrow=TRUE)

```


## Restructure the data
```{r}
dotLineSeq <- seq(1,length(tournamentData),3)
#Remove all --- line 
linn1<-tournamentData[-dotLineSeq]

# Read all the alternate lines 

#Read player names
playerInfo <- linn1[seq(3,length(linn1),2)]

# Read state and pre-tournament ratings
playerDetails <- linn1[seq(4,length(linn1),2)]

#combile the player information and state information into single vector
playerInfo<-paste0(playerInfo,playerDetails)

playerInfo


```

## Create a dataframe with player information

```{r}
# Create an Empty data frame 
playerInfoDf <- data.frame(data1 = character(0),Name = character(0),data3= character(0),Round1= character(0),Round2= character(0),Round3= character(0),Round4= character(0),Round5= character(0),Round6= character(0),Round7= character(0),state= character(0),data12= character(0))

# Split the string by | and populate records in data frame (only first 12 columns)
i<- 1
while (i <= length(playerInfo)){
    playerInfoDf[nrow(playerInfoDf) + 1, ] <- str_trim(strsplit(playerInfo[i],"|",fixed=TRUE)[[1]])[1:12]
    i <- i+1
}

head(playerInfoDf)

```



## Create final dataframe for CSV
```{r}
#Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents 

finalResultCSVFrame <- data.frame(Name = character(0),State = character(0),TotalPoints= character(0),PreRating= character(0),AvgRatingOfOpp= character(0))

#iterate each record
for(i in 1:nrow(playerInfoDf)){
  value3 <- 0
  cnt <-0
   # extract opponent id for each round  (1 to 7 ) -from column 4,10
   for (j in 4:10){
     
     #split each round column by space
     roundVals <- strsplit(playerInfoDf[i,j], "\\s+")
     
     if(roundVals[[1]][1]== "W" ||  roundVals[[1]][1]== "L"||  roundVals[[1]][1]== "D"){
       
       #using the row id ,find opponent Pre-rating column
       value2 <- playerInfoDf[roundVals[[1]][2],12]
       
       #find the 2nd number ,ie number after R: ,and sum it
       value3 <- value3 + as.numeric(regmatches(value2 , gregexpr("[[:digit:]]+", value2 ))[[1]][2])
       cnt <- cnt +1
     }
     
   }
   #find the mean value and round it
   meanValue <- round(value3/cnt)
 
  #find the player pre-rating info  
  playerPreValue <-  as.numeric(regmatches(playerInfoDf[i,12] , gregexpr("[[:digit:]]+", playerInfoDf[i,12] ))[[1]][2])
  
  finalResultCSVFrame[nrow(finalResultCSVFrame) + 1, ] <-c(playerInfoDf[i,'Name'],playerInfoDf[i,'state'],playerInfoDf[i,'data3'],playerPreValue,meanValue)
}

finalResultCSVFrame

#write the data into CSV file
write.csv(finalResultCSVFrame,"playerInfo.csv")
```


### Conclusion

The application read the text input file, transform the data into required output format and finally write the data frame into a CSV file (playerInfo.csv)  



