---
title: "Week5 Assignment"
output: html_document
---
#### ***Rathish Parayil Sasidharan***


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The chart above describes arrival delays for two airlines across five destinations.  \

Your task is to: 
(1) Create a .CSV file  that includes all of the information above.  \
    You’re encouraged to use a “wide” structure similar to how the information appears above, \ 
    so that you can practice tidying and transformations as described below. \
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as \ 
   needed to tidy and transform your data.\   
(3) Perform analysis to compare the arrival delays for the two airlines.\ 
(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data\     cleanup work, analysis, and conclusions.\
Please include in your homework submission: \
The URL to the .Rmd file in your GitHub repository. and 
The URL for your rpubs.com web page

### Prerequisites
Load required packages
```{r message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
```

## Read the airline data

```{r message=FALSE}
airlinesUntidyDf<-read.csv("https://raw.githubusercontent.com/rathish-ps/Data607-Assignment/main/airlines.csv")
airlinesUntidyDf

```


## Clean up the data

Transform the data from wide format to long format

```{r message=FALSE}
#Move the destination columns (3 to 7) from the untidy data frame  into single destination column
#wide data into longer format and then widen the Arrival Status into separate columns

airlinesDf <- gather(airlinesUntidyDf,Destination, delays, 3:7) %>%spread(2,4)
airlinesDf
```


## Data transformation

Add additional columns for total ,  percentages for On time and delays

```{r message=FALSE}

#Add additional columns for total flights , on time and delayed percentages 
airlinesTfDf <- airlinesDf %>% 
        mutate(total = delayed + `on time`,prcntOnTime = round(100 * `on time` / total, 2), prcntDelayed = round(delayed / total * 100, 2)) 


airlinesTfDf

#summary(airlinesTfDf)

#Create a summary record for each airline
 airlineSummaryDf <- airlinesTfDf %>% 
  group_by(Airline) %>% 
  summarise(totalDelayed = sum(delayed), totalOnTime = sum(`on time`), totalFlights = sum(total)) 
 
# Add additional columns for percentage - on-time and delayed flights 
 airlineSummaryDf<- airlineSummaryDf %>%  mutate(percentOntime = round(totalOnTime/totalFlights * 100, 2),percentDelayed = round(totalDelayed/totalFlights * 100, 2))
airlineSummaryDf
```


## Data visualisation
Bar plot for each destination by Airline
```{r message=FALSE}

#bar plot for ontime flights
plt <- ggplot(airlinesTfDf, aes(y=prcntOnTime , x = Airline, color = Airline, fill = Airline)) + 
    geom_bar( stat = "identity") +
    geom_text(aes(label = prcntOnTime ), vjust = -.35) +
     ylim(0, 95) +
    facet_wrap(~Destination) +
    ylab("Percentage - On time Flights") +
    ggtitle("Percentage of On time Flights / Airline / Destination")
    
plt


# bar plot for delayed flights
plt1 <- ggplot(airlinesTfDf, aes(y=prcntDelayed, x = Airline, color = Airline, fill = Airline)) + 
    geom_bar( stat = "identity") +
    geom_text(aes(label = prcntDelayed), vjust = -.35) +
    ylim(0, 30) +
    facet_wrap(~Destination) +
    ylab("Percentage - Delayed Flights") +
    ggtitle("Percentage of Delayed Flights / Airline / Destination")
    
plt1




```


## Summary

It is clear from the summary data that 'ALASKA' airline has more percentage delay than 'AM WEST' airline ,even though
'AM WEST' has more percentage delay for each destination. This is mostly because of the number of flights operated by each airline to different destinations are different. 





