---
title: "Data 607 Week 3"
output: html_document
---
#### ***Rathish Parayil Sasidharan***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to the problems below. 

## Problem 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset 
[https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

### Prerequisites
Load required packages
```{r message=FALSE}
library("stringr")

```


Read the CSV file and grep for "DATA" or "STATISTICS"
```{r}

#Read the file 

col_majors_ds<-read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

col_majors_ds[grep("DATA", col_majors_ds$Major,ignore.case ="True"),]
col_majors_ds[grep("STATISTICS", col_majors_ds$Major,ignore.case ="True"),]
```

## Problem 2 
2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

```{r}

input_str <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

extracted_str <- str_extract_all(input_str, pattern = "\"([:alpha:]+.[:alpha:]+)\"")

final_str <- str_replace_all(extracted_str[[1]], "\"", "")

final_str

```

## Problem 3 

Describe, in words, what these expressions will match:

(.)\1\1 

**Ans -**\
Pattern for matching any single character, followed by two repetitions (3 repeating characters)

eg : aaa, 111
```{r}
testData <- c("111")
str_extract_all(testData , regex("(.)\\1\\1"))
 
```


"(.)(.)\\2\\1"

**Ans **
I assume \\ and \\\\ (escaped \\) are the same ,even though they are not literally same.
In R , \\ has to be escaped (\\\\). 
But with regard to a regular expression context  \\\\1 means character \\ followed by group 1 and \\1 mean the group 1.

Two consecutive characters are repeated but in reverse order
eg: 1221 or abba 

```{r}
testData <- c("abba")
str_extract_all(testData , regex("(.)(.)\\2\\1"))
 
```

(..)\1

**Ans **
Two consecutive characters are repeated
eg : abab, 1212

```{r}
testData <- c("abab")
str_extract_all(testData , regex("(..)\\1"))
 
```


"(.).\\1.\\1"

**Ans **
In a string ,if there is a 5 consecutive character with 3 are repeating which are in 1,3,5 character positions.

eg :12141 ,abaca


```{r}
testData <- c("12141")
str_extract_all(testData , regex("(.).\\1.\\1"))
 
```


"(.)(.)(.).*\\3\\2\\1"

***Ans ** \

a set of characters  with first 3 and last 3 characters are same ,but in reverse order with any number of characters in between them \

eg : abc12345cba \

{firstcharacter}{secondcharacter}{thirdcharacter}{any number of characters}{group3}{group2}{group1}\

```{r}
testData <- c("abc12345cba")
str_extract_all(testData , regex("(.)(.)(.).*\\3\\2\\1"))
```

## Problem 4 - Construct regular expressions to match words that:

Start and end with the same character.

**Ans**\
"(.).*\\1"


Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

**Ans **\

".*([a-z][a-z]).*\1.*"  - without escape

".\*([a-z][a-z]).\*\1.*" --with escape

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

**Ans** \

".*([a-z]).*\1.*\1.*" - without escape

".\*([a-z]).\*\1.\*\1.\*" --with escape



