---
title: "Week9Assignment"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true

---
#### ***Rathish Parayil Sasidharan***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The New York Times web site provides a rich set of APIs, as described here:  [https://developer.nytimes.com/apis] You’ll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame. 


## Movie Reviews API

For this assignment we will be using API to Get movie reviews that are critics' picks:

/reviews/picks.json



## Prerequisite
Create a developer account on the developer.nytimes.com and register an application and sign up for API Key 

Load the required libraries
```{r , message=FALSE}
library('jsonlite')
library('httr')
library('tidyverse')
```


## Read the API Key

Read the API key provided by New York Time while registering the app
```{r}
pwfile <- "apiKey.txt"
if ( file.exists(pwfile) )
{
  apiKeyValue <- readLines(pwfile)
}else 
{
  apiKeyValue <- rstudioapi::askForPassword( prompt = "Please enter API key:" ) 
}

```


## Construct URL by appending the API Key 
Append the API key to the base url

```{r}
nytmoviePickBaseUrl <- 'https://api.nytimes.com/svc/movies/v2/reviews/picks.json'
finalUrl <- paste0(nytmoviePickBaseUrl,"?api-key=", apiKeyValue)

```

## Read the data and display the results
```{r}
moviePicksJson <-read_json(finalUrl, simplifyVector = TRUE)


moviePicksJson$status

#Check the http status code
if(moviePicksJson$status =='OK'){
     #Select first 8 columns and create a new data frame
     moviePicksJsonDf <- moviePicksJson$results %>% select(1:8)
     DT::datatable(moviePicksJsonDf)
}
```

## Summary
Successfully retrieved the top movie reviews that are critics picks by using the developer API
Converted the json data into data frame
