---
title: "Week10 Assignment"
author: "Rathish Sasidharan"
date: "4/16/2021"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary \ example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code. \ 
You’re then  asked to extend the code in two ways: \

Work with a different corpus of your choosing, and \
Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research). \
As usual, please submit links to both an .Rmd file posted in your GitHub repository and to your code on rpubs.com.  \
You make work on a small team on this assignment.

```{r}
# Load library

library(stringr)
library(tidytext)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(dplyr)
library(textdata)
library(wordcloud)
library(janeaustenr)

library(reshape2)
library(ggwordcloud)
library(wordcloud)
library(gutenbergr)

```



```{r}

tidy_book <- function(author) {
  author %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words)
}


facet_bar <- function(df, y, x, by, nrow = 2, ncol = 2, scales = "free") {
  mapping <- aes(y = reorder_within({{ y }}, {{ x }}, {{ by }}), 
                 x = {{ x }}, 
                 fill = {{ by }})
  
  facet <- facet_wrap(vars({{ by }}), 
                      nrow = nrow, 
                      ncol = ncol,
                      scales = scales) 
  
  ggplot(df, mapping = mapping) + 
    geom_col(show.legend = FALSE) + 
    scale_y_reordered() + 
    facet + 
    ylab("")
} 
```

## 2.2 Sentiment analysis with inner join

```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  mutate(index = linenumber %/% 80) %>% 
  count(book, index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = list(n = 0)) %>%
  mutate(sentiment = positive - negative)

jane_austen_sentiment


ggplot(jane_austen_sentiment) + 
  geom_col(aes(index, sentiment, fill = book), show.legend = F) + 
  facet_wrap( ~ book, ncol = 2, scales = "free_x") 

```

## 2.3 Comparing 3 different dictionaries

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  mutate(index = linenumber %/% 80) %>% 
  count(book, index, wt = value, name = "value") %>%
  mutate(dict = "afinn") %>% 
  select(index, value, dict)


bing <- pride_prejudice %>% 
  inner_join(get_sentiments("bing")) %>% 
  mutate(index = linenumber %/% 80) %>% 
  count(index, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = list(n = 0)) %>% 
  mutate(value = positive - negative,
         dict = "bing") %>%  
  select(index, value, dict)


nrc <- pride_prejudice %>% 
  inner_join(get_sentiments("nrc")) %>%
  filter(sentiment %in% c("positive", "negative")) %>% 
  mutate(index = linenumber %/% 80) %>% 
  count(index, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = list(n = 0)) %>% 
  mutate(value = positive - negative,
         dict = "nrc") %>% 
  select(index, value, dict)

bind_rows(afinn, bing, nrc) %>% 
  ggplot() + 
  geom_col(aes(index, value, fill = dict), show.legend = FALSE) + 
  facet_wrap(~ dict, nrow = 3)


get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)

```

## 2.4 Most common positive and negative words

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

custom_stop_words <- tibble(word = c("miss"), lexicon = c("custom")) %>% 
  bind_rows(stop_words)

bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>% 
  anti_join(custom_stop_words) %>%
  group_by(sentiment) %>%
  count(word, sentiment, sort = T) %>% 
  ungroup()

bing_word_counts %>%
  group_by(sentiment) %>% 
  top_n(10) %>%
  ungroup()  %>% 
  facet_bar(y = word, x = n, by = sentiment, nrow = 1) + 
  labs(title = "Top 10 words of sentiment in Jane Austen's books") 

```

## 2.5 Wordclouds

```{r}

wordcloud_df <-tidy_books %>% 
  anti_join(custom_stop_words) %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(sentiment, word, sort = T) %>% 
  top_n(200)

wordcloud_df %>%
  ggplot() + 
  geom_text_wordcloud_area(aes(label = word, size = n)) +
  scale_size_area(max_size = 15)

wordcloud_df %>%
  ggplot() + 
  geom_text_wordcloud_area(aes(label = word, size = n), shape = "star") + 
  scale_size_area(max_size = 15)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)


```

## 2.6 Units other than words

```{r}
PandP_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

PandP_sentences

tibble(text = prideprejudice) %>% 
  mutate(text = iconv(text, to = "ASCII")) %>% 
  unnest_tokens(sentence, text, token = "sentences")

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters

tidy_books %>%
  distinct(book, chapter)

tidy_books %>% 
  group_by(book, chapter) %>% 
  summarize(str_c(word, collapse = " "))

bing_negative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

chapter_words <- tidy_books %>% 
  count(book, chapter)

tidy_books %>%
  semi_join(bing_negative) %>%
  count(book, chapter, name = "negative_words") %>% 
  left_join(chapter_words) %>%
  mutate(ratio = negative_words / n) %>%
  filter(chapter != 0) %>%
  group_by(book) %>% 
  top_n(1) 

```


## Work with a different corpus

```{r warning = FALSE}
# filter a book  by author title language and copyright
aesopMtdt <-gutenberg_metadata %>% 
    filter(author == "Aesop",
           title == "Aesop's Fables",language == "en",!grepl('Copyrighted', rights))

# load the book using book id
aesopBook <- gutenberg_download(aesopMtdt$gutenberg_author_id[1])

#mutate to add author
aesopBookMutated <- aesopBook %>% 
  select(-gutenberg_id) %>% 
  mutate(author = "Aesop")

#aesopBookTidy<-tidy_book(aesopBookMutated)

# identify the tokens and remove stop words
aesopBookTidy<-aesopBookMutated %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words)
 


```


## sentiment lexicon loughran
```{r warning = FALSE}
#load the sentiment lexicon loughran
get_sentiments("loughran")

# join the sentiments with tokenized books with count for each sentiment  
aesopBook_lexi <- aesopBookTidy %>%
inner_join(get_sentiments("loughran")) %>%
count(word, sentiment, sort = TRUE) %>%
ungroup()

```

## Identify top 10 sentiments
```{r warning = FALSE}
aesopBook_lexi %>%
group_by(sentiment) %>%
top_n(10) %>%
ungroup() %>%mutate(word = reorder(word, n)) 

```


## Identify positive and negative sentiments

```{r warning = FALSE}
#filter out positive and negative sentiments
aesopBook_positive_negative<-aesopBook_lexi %>%filter(sentiment =='positive' | sentiment =='negative')

aesopBook_nrc_lexi <- aesopBookTidy %>%
inner_join(get_sentiments("nrc")) %>%
count(word, sentiment, sort = TRUE) %>%
ungroup()

aesopBook_positive_negative_nrc<-aesopBook_nrc_lexi %>%filter(sentiment =='positive' | sentiment =='negative')


```


## Plot for positive and negative sentiments for loughran

```{r warning = FALSE}

aesopBook_positive_negative %>%
     filter(n > 50) %>%
     mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
     mutate(word = reorder(word, n)) %>%
     ggplot(aes(word, n, fill = sentiment))+
     geom_col() +
     coord_flip() +
     labs(y = "Sentiment Count")
```

## Plot for positive and negative sentiments for nrc

```{r warning = FALSE}
aesopBook_positive_negative_nrc %>%
     filter(n > 100) %>%
     mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
     mutate(word = reorder(word, n)) %>%
     ggplot(aes(word, n, fill = sentiment))+
     geom_col() +
     coord_flip() +
     labs(y = "Sentiment Count")
```


## word clout

```{r warning = FALSE}
# Word clouts
aesopBookTidy %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))


# sentiment analysis to tag different sentiments
 aesopBookTidy %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```


## Conclusion

Identified a different corpus and completed sentimental analysis using loughran lexicon.
Also performed a sentimental comparison of the words extracted from the same book using loughran and nrc lexicons.
It is clear that different lexicon dictionaries are having different set of sentimental words 
